---
title: 'Results: Compare and Contrast'
date: "3/3/2020"
output:
  pdf_document: default
  #html_document: default
---

# Results: Compare and Contrast

This next section is my attempt to compare and contrast the ML Algorithms prepared in this  report.

As you will remember, I have studied 6 different M.L. algorithms using protein amino acid percent composition data from two classes. Class number 1 is my positive control which is a set of Myoglobin proteins, while the second class is a control group of human proteins that do not have Fe binding centers.

| Group     |  Class   | N of Class | Range of Groups |
| :-------- | :------: | :--------: | --------------: |
| Controls  | 0 or (-) |    1217    |    1, ..., 1217 |
| Myoglobin | 1 or (+) |    1124    | 1218, ..., 2341 |

The Six M.L Algorithms consist of:

| Name                             |         Type |    Output Used For Graphing |
| :------------------------------- | -----------: | :-------------------------: |
| Principal Component Analysis     | Unsupervised | Anomalies > Abs($3 \sigma$) |
| Logistic Regression              |   Supervised |                     FP & FN |
| SVM-linear                       |   Supervised |                     FP & FN |
| SVM-polynomial kernel            |   Supervised |                     FP & FN |
| SVM-radial basis function kernel |   Supervised |                     FP & FN |
| Neural Network                   |   Supervised |                     FP & FN |


=========================

## Scatter Plots of Anomalies Vs. FP & FN Outputs

*To obtain False-Positive (fp) from sets:* {-}

1. False-Postives $~~~\stackrel{\mathrm{def}}{=}~$ {obs = 0 $~\wedge~$ pred = 1}  
2. False-Negatives $~\stackrel{\mathrm{def}}{=}~$ {obs = 1 $~\wedge~$ pred = 0}  


### Anomalies Inner Joined with PC

```{r message=FALSE, warning=FALSE}
## Load Libraries
Libraries = c("knitr", "readr")

for(p in Libraries){
    library(p, character.only = TRUE)
}
opts_chunk$set(cache = TRUE, fig.align = "center")
```

## Prepare PCA: PC1 and PC2 for all 2340 proteins

```{r message=FALSE, warning=FALSE}
norm_c_m_20aa <- read_csv("./00-data/03-ml_results/norm_c_m_20aa.csv")

pca_values <- prcomp(norm_c_m_20aa)

row_pc12 <- cbind(rowNum = 1:2340, PC1 = pca_values$x[,1], PC2 = pca_values$x[,2])
# dim(row_pc12)
```

```{r include=FALSE}
write.table(row_pc12,
            file = "./00-data/05-joins_plots/row_pc12.txt",
            sep = ",",
            row.names = F)
```

Load Files & change column names
```{r warning=FALSE, include=FALSE}
# Not elegant but fast 4 me to code ;)
logit_nums <- read_csv("./00-data/04-sort_unique_outliers/logit_nums.csv")
pca_outliers <- read_csv("./00-data/04-sort_unique_outliers/pca_outliers.csv")
# rf_nums <- read_csv("./00-data/04-sort_unique_outliers/rf_nums.csv")
svm_lin_nums <- read_csv("./00-data/04-sort_unique_outliers/svm_lin_nums.csv")
svm_poly_nums <- read_csv("./00-data/04-sort_unique_outliers/svm_poly_nums.csv")
svm_rbf_nums <- read_csv("./00-data/04-sort_unique_outliers/svm_rbf_nums.csv")

colnames(logit_nums)[1] = "rowNum" 
colnames(pca_outliers)[1] = "rowNum" 
# colnames(rf_nums)[1] = "rowNum" 
colnames(svm_lin_nums)[1] = "rowNum" 
colnames(svm_poly_nums)[1] = "rowNum" 
colnames(svm_rbf_nums)[1] = "rowNum" 
```

Inner-join (merge) with PC 1&2 (row_pc12)
```{r message=FALSE, warning=FALSE}
logit_pc    <- merge(x = logit_nums,    y = row_pc12, by = "rowNum")
pca_pc      <- merge(x = pca_outliers,  y = row_pc12, by = "rowNum")
# rf_pc       <- merge(x = rf_nums,       y = row_pc12, by = "rowNum")
svm_lin_pc  <- merge(x = svm_lin_nums,  y = row_pc12, by = "rowNum")
svm_poly_pc <- merge(x = svm_poly_nums, y = row_pc12, by = "rowNum")
svm_rbf_pc  <- merge(x = svm_rbf_nums,  y = row_pc12, by = "rowNum")
```

## Generate FP/FN & PCA-Anomalies on PCA Plots

### PCA-Anomalies Plot
```{r echo=FALSE, fig.height=7, fig.width=7}
plot(x = row_pc12[, 2], 
     y = row_pc12[, 3], 
     main = "PCA-Anomalies On PC 1 & 2 Axes",
     sub = "(PCA-Anomalies pts are overlayed on PCA)",
     xlab = "PC1",
     ylab = "PC2",
     pch = 16,
     cex = 0.8,
     col = "gray87")
opar <- par(new = TRUE, mar = c(0,0,0,0))
text(pca_pc[, 2], 
     pca_pc[, 3], 
     labels = pca_pc[, 1], 
     col = "red",
     cex = 1.4, 
     pos = 1)
```

### Logit Plot
```{r echo=FALSE, fig.height=7, fig.width=7}
plot(x = row_pc12[, 2], 
     y = row_pc12[, 3], 
     main = "Logit FP/FN On PC 1 & 2 Axes",
     sub = "(Logit pts are overlayed on PCA)",
     xlab = "PC1",
     ylab = "PC2",
     pch = 16,
     cex = 0.8,
     col = "gray87")
opar <- par(new = TRUE, mar = c(0,0,0,0))
text(logit_pc[, 2], 
     logit_pc[, 3], 
     labels = logit_pc[, 1], 
     col = "coral",
     cex = 1.4, 
     pos = 1)
```

### SVM-Linear Plot
```{r echo=FALSE, fig.height=7, fig.width=7}
plot(x = row_pc12[, 2], 
     y = row_pc12[, 3], 
     main = "SVM-Linear FP/FN On PC 1 & 2 Axes",
     sub = "(SVM-Linear pts are overlayed on PCA)",
     xlab = "PC1",
     ylab = "PC2",
     pch = 16,
     cex = 0.8,
     col = "gray87")
opar <- par(new = TRUE, mar = c(0,0,0,0))
text(svm_lin_pc[, 2], 
     svm_lin_pc[, 3], 
     labels = svm_lin_pc[, 1], 
     col = "darkgoldenrod3",
     cex = 1.4, 
     pos = 1)
```

### SVM-Polynomial Plot
```{r echo=FALSE, fig.height=7, fig.width=7}
plot(x = row_pc12[, 2], 
     y = row_pc12[, 3], 
     main = "SVM-Polynomial FP/FN On PC 1 & 2 Axes",
     sub = "(SVM-Polynomial pts are overlayed on PCA)",
     xlab = "PC1",
     ylab = "PC2",
     pch = 16,
     cex = 0.8,
     col = "gray87")
opar <- par(new = TRUE, mar = c(0,0,0,0))
text(svm_poly_pc[, 2], 
     svm_poly_pc[, 3], 
     labels = svm_poly_pc[, 1], 
     col = "deepskyblue3",
     cex = 1.4, 
     pos = 1)
```

### SVM-Radial Basis Function Plot
```{r echo=FALSE, fig.height=7, fig.width=7}
plot(x = row_pc12[, 2], 
     y = row_pc12[, 3], 
     main = "SVM-RBF FP/FN On PC 1 & 2 Axes",
     sub = "(SVM-RBF pts are overlayed on PCA)",
     xlab = "PC1",
     ylab = "PC2",
     pch = 16,
     cex = 0.8,
     col = "gray87")
opar <- par(new = TRUE, mar = c(0,0,0,0))
text(svm_rbf_pc[, 2], 
     svm_rbf_pc[, 3], 
     labels = svm_rbf_pc[, 1], 
     col = "green4",
     cex = 1.4, 
     pos = 1)
```


### Neural Network Function Plot







## Statistical Learning Method Vs Total Number of FP/FN

| Statistical Method            | Total Number Produced | Unique | Total/Unique |
| :---------------------------- | :-------------------: | :----: | :----------: |
| Principal Componnent Analysis |          461          |  460   |    1.002     |
| Logit                         |          537          |  119   |     4.51     |
| SVM Linear                    |          496          |  125   |     3.97     |
| SVM Polynomial                |          278          |   70   |     3.97     |
| SVM Radial Basis Function     |          244          |   58   |     4.21     |
| Random Forest                 |          190          |   46   |     4.13     |
| Deep Learning                 |          347          |  133   |     2.61     |















=======================

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.align="center")
```

```{r message=FALSE, warning=FALSE}
# Load Libraries
Libraries <- c("doMC", "knitr", "readr", "caret", "nnet", "caretEnsemble", "e1071", "kernlab")
for (p in Libraries) {  
    library(p, character.only = TRUE)
}
```


```{r 611}
# Import data & data handling
c_m_TRANSFORMED <- read_csv("./00-data/02-aac_dpc_values/c_m_TRANSFORMED.csv",
                            col_types = cols(Class = col_factor(levels = c("0", "1")),
                                             PID = col_skip(),
                                             TotalAA = col_skip()))
```

```{r 612}
# Partition data into training and testing sets
set.seed(1000)
index <- createDataPartition(c_m_TRANSFORMED$Class, p = 0.8, list = FALSE)

training_set <- c_m_TRANSFORMED[ index,]
test_set     <- c_m_TRANSFORMED[-index,]

Class_test <- as.factor(test_set$Class)
```

### Stacking Algorithms - Run multiple algorithms in one call.
```{r include=FALSE}
# new trainControl object with index specified
trainControl <- trainControl(method = "repeatedcv",
                             number = 10,
                             index = createFolds(training_set$Class, 10),
                             repeats = 5,
                             savePredictions = "all",
                             search = "random")

algorithmList <- c('glm', 'nnet', 'svmLinear', 'svmPoly', 'svmRadialCost')

set.seed(100)
registerDoMC(cores = 3) # Start multi-processor mode
start_time <- Sys.time() # Start timer

models <- caretList(Class ~ .,
                    data = training_set,
                    trControl = trainControl,
                    methodList = algorithmList)

results <- resamples(models)
```

Plot the resamples output to compare the models.
```{r}
# Box plots to compare models
scales <- list(x = list(relation = "free"),
               y = list(relation = "free"))
bwplot(results, scales = scales)
```















