 ---
output:
  pdf_document: default
  html_document: default
---


## EDA, Analysis of RAW data

Raw data is considered: `./00-data/02-aac_dpc_values/c_m_RAW_AAC.csv`

```{r 21, message=FALSE}
# Import libraries, NO , 
library(easypackages)
libraries("knitr", "doMC", "readr", "RColorBrewer", "corrplot", "Boruta", "kableExtra")
```

```{r 22, include=FALSE}
opts_chunk$set(cache = TRUE, fig.pos = "center")
```

```{r 23, message=FALSE, warning=FALSE}
# Import RAW data
c_m_RAW_AAC <- read_csv("./00-data/02-aac_dpc_values/c_m_RAW_AAC.csv")
Class <- as.factor(c_m_RAW_AAC$Class)
```

### Visually inspect RAW data files

1. Use the command-line interface followed by the command `less.`
2. Check for binary instead of ASCII and bad Unicode.

### Inspect RAW dataframe structure, `str()` 

```{r 24, echo=FALSE}
str(c_m_RAW_AAC)
```



### Check RAW data `head` & `tail`

```{r 25}
head(c_m_RAW_AAC, n = 2)
```
```{r 26}
tail(c_m_RAW_AAC, n = 2)
```

### Check RAW data types

```{r 27, echo=TRUE}
is.data.frame(c_m_RAW_AAC)
class(c_m_RAW_AAC$Class) # Col 1
class(c_m_RAW_AAC$TotalAA) # Col 2
class(c_m_RAW_AAC$PID) # Col 3
class(c_m_RAW_AAC$A) # Col 4
```

### Check RAW dataframe dimensions

```{r 28}
dim(c_m_RAW_AAC)
```

### Check RAW for missing values

- **No missing values found.**
```{r 29}
apply(is.na(c_m_RAW_AAC), 2, which)

# sapply(c_m_RAW_AAC, function(x) sum(is.na(x))) # Sum up NA by columns
# c_m_RAW_AAC[rowSums(is.na(c_m_RAW_AAC)) != 0,] # Show rows where NA's is not zero
```

### Number of polypeptides per Class:

- Class 0 = Control,
- Class 1 = Myoglobin
```{r 210, echo=FALSE}
class_table <- table(c_m_RAW_AAC$Class)
class_table
```

### Numerical summary of RAW data

```{r 211, echo=FALSE}
summary(c_m_RAW_AAC)
```

### Visualize Descriptive Statistics using RAW Data

Formulas for mean:
\begin{equation} 
E[X] = \sum_{i=1}^n x_i p_i ~~; ~~~~~~ \bar x = \frac {1}{n} \sum_{i=1}^n x_i
\end{equation}

### Scatter plot of means of *Myoglobin-Control* amino acid composition of RAW Data

- This Scatter-plot shows the means for each feature (column-means) in the dataset. The means represent the ungrouped or total of all proteins (where n = 2340) versus AA type.

```{r 212,echo=FALSE,out.width='60%',fig.align='center'}
AA_ave <- colMeans(c_m_RAW_AAC[, 4:23])
plot(AA_ave,
     main = "RAW Data: Column-Means of % Composition Vs Amino Acid",
     ylab = "% Composition",
     xlab = "Amino Acid",
     ylim = c(0, 0.1),
     type = "b",
     xaxt = "n")
axis(1, at = 1:20, labels = names(c_m_RAW_AAC[, 4:23]))
```

```{r 213, cache=TRUE, include=FALSE}
# * Pseudo-code for Data manipulation for Grouped Bar charts
# + A-1. Subset 7 protein groups, [Control:Ctrl, Myoglobin:Mgb] & Grand-Mean of both sets
# + A-2. Determine column means for each protein class
# + A-3. Calculate percentage values
# + A-4. Produce Grouped Bar Plot

# A-1
ctrl_set <- c_m_RAW_AAC[which(c_m_RAW_AAC$Class == "0"), ]
mgb_set  <- c_m_RAW_AAC[which(c_m_RAW_AAC$Class == "1"), ]

# A-2
ctrl_means <- apply(ctrl_set[, 4:23], 2, mean)
mgb_means  <- apply(mgb_set[, 4:23], 2, mean)
grand_mean <- apply(c_m_RAW_AAC[, 4:23], 2, mean)

# A-3
data <- data.frame(ctrl_means, mgb_means, grand_mean)
percent_aa <- as.matrix(t(100 * data))
```

```{r 214,echo=TRUE,eval=FALSE,message=FALSE,warning=FALSE}
# A-4
### Grouped barchart of amino acid vs. protein category
barplot(percent_aa,
        main = "Mean % A.A.Composition Of 3 Protein Groupings",
        ylab = "% AA Composition",
        ylim = c(0, 12),
        col = colorRampPalette(brewer.pal(4, "Blues"))(3),
        legend = T,
        beside = T)
```

### Means of percent amino acid composition of control & myoglobin categories, RAW data

```{r 215,echo=FALSE,message=FALSE,warning=FALSE,cache=TRUE}
data2 <- data.frame(ctrl_means, mgb_means)
percent_aa2 <- as.matrix(t(100 * data2))

barplot(percent_aa2,
        ylim = c(0, 12),
        main = "Mean % A.A.Composition Of Control & Myoglobin",
        ylab = "% AA Composition",
        col = colorRampPalette(brewer.pal(4, "Blues"))(2),
        legend = T,
        beside = T)
```

### Boxplots of grand-means of overall amino acid composition, RAW data

```{r 216,echo=FALSE,message=FALSE,warning=FALSE}
boxplot(c_m_RAW_AAC[, 4:23],
        main = "Boxplots: All; % Composition Vs Amino Acid",
        ylab = "% AAC",
        xlab = "Amino Acid",
        las = 1)
```

### Boxplots of amino acid compositions for control (only), RAW data

```{r 217,echo=FALSE}
boxplot(ctrl_set[, 4:23],
        main = "Boxplots: Controls; % AAC Vs Amino Acid",
        ylab = "% AAC",
        xlab = "Amino Acid",
        las = 1)
```

### Boxplots of amino acid compositions for myoglobin (only), RAW data

```{r 218,echo=FALSE}
boxplot(mgb_set[, 4:23],
        ylim = c(0, 0.5),
        main = "Boxplot: Myoglobin; % AAC Vs Amino Acid",
        ylab = "% AAC",
        xlab = "Amino Acid",
        las = 1)
```

```{r 219,echo=FALSE}
par(mfrow = c(1, 2))

boxplot(ctrl_set[, 4:23],
        ylim = c(0, 0.3),
        main = "Boxplots: Controls",
        ylab = "% AAC",
        xlab = "Amino Acid",
        las = 1)

boxplot(mgb_set[, 4:23],
        ylim = c(0, 0.3),
        main = "Boxplot: Myoglobin",
        xlab = "Amino Acid",
        las = 1)
```

### Boxplots Of Length Of Polypeptides For RAW Data 

```{r 220,echo=FALSE}
ctrl_totalaa  <- ctrl_set[, 2]
mgb_totalaa   <- mgb_set[, 2]
grand_totalaa <- c_m_RAW_AAC[, 2]

data <- c(ctrl_totalaa, mgb_totalaa, grand_totalaa)

boxplot(data,
        ylim = c(0, 5000),
        main = "RAW Data: Length of Polypeptides Vs Control, Myoglobin & Combined",
        ylab = "Length of Polypeptides",
        xaxt = "n",
        las = 2)
axis(1, at = 1:3, labels = c("Control", "Myoglobin", "Combined"))
```

### Plot Coefficient Of Variance For RAW Data

Standard deviations are sensitive to scale. Therefore I compare the normalized standard deviations. This normalized standard deviation is more commonly called the coefficient of variation (CV).

\begin{equation} 
CV = \frac {\sigma (x)} {E [|x|]} ~~~ where ~~~ \sigma(x) \equiv \sqrt{ E[x - \mu]^2 }
\end{equation}

\begin{equation} 
CV ~~=~~ \frac{1}{\bar x} \cdot \sqrt{ \frac{1}{n-1} \sum^n_{i=1} (x_i - \bar x)^2}
\end{equation}

```{r 221, echo=FALSE}
AA_var_norm <- (apply(c_m_RAW_AAC[, 4:23], 2, sd)) / AA_ave
plot(AA_var_norm,
     main = "Plot of Coefficient Of Variance (CV), RAW Data",
     sub = "(Note: Two largest values shown in red.)",
     ylab = "Coefficient Of Variance (CV)",
     xlab = "Amino Acid",
     ylim = c(0, 1.5),
     type = "b",
     xaxt = "n")
axis(1, at = 1:20, labels = names(c_m_RAW_AAC[, 4:23]))
text(x = 2, y = 1.4, label = " C=1.24", col = "red")
text(x = 19, y = 1.1, label = "W=0.946", col = "red")
```

```{r 222}
AA_var_norm
```



### Skewness of distributions, RAW data

\begin{equation} 
Skewness ~= E\left[ \left( \frac{X - \mu}{\sigma(x)} \right)^3 \right] ~~~~ where ~~~~ \sigma(x) \equiv \sqrt{ E[x - \mu]^2 }
\end{equation}

\begin{equation} 
Skewness ~= \frac { \frac{1}{n} \sum^n_{i=1} (x_i - \bar x)^3 } { \left( \sqrt{ \frac{1}{n-1} \sum^n_{i=1} (x_i - \bar x)^2 } \right) ^ {3}}
\end{equation}

Skewness values for each A.A. are determined in totality.

```{r 223, echo=FALSE}
AA_skewness <- (apply(c_m_RAW_AAC[, 4:23], 2, e1071::skewness))
plot(AA_skewness,
     main = "Plot of Skewness Vs Amino Acids, RAW Data",
     ylab = "Skewness",
     xlab = "Amino Acid",
     type = "b",
     ylim = c(-0.5, 3),
     xaxt = "n")
axis(1, at = 1:20, labels = names(c_m_RAW_AAC[, 4:23]))
abline(h = 2.0, col = "red")
text(x = 2, y = 2.8, label = " C=2.5", col = "red")
text(x = 5, y = 2.4, label = "F=2.1", col = "red")
text(x = 8, y = 2.4, label = "I=2.2", col = "red")
```

```{r}
AA_skewness
```

### Determine coefficients of correlation, RAW data

An easily interpretable test is a correlation 2D-plot for investigating multicollinearity or feature reduction. Fewer attributes "means decreased computational time and complexity. Secondly, if two predictors are highly correlated, this implies that they measure the same underlying information. Removing one should not compromise the performance of the model and might lead to a more parsimonious and interpretable model. Third, some models can be crippled by predictors with degenerate distributions." [^27]

[^27]:Max Kuhn and Kjell Johnson, Applied Predictive Modeling, Springer Publishing, 2018, P.43

Pearson's correlation coefficient: 
\begin{equation} 
\rho_{x,y} = \frac {E \left[(X - \mu_x)(X - \mu_y) \right]} {\sigma_x \sigma_y}
\end{equation}

\begin{equation} 
r_{xy} = \frac {\sum^n_{i=1} (x_i - \bar x)(y_1 - \bar y)} { {\sqrt {\sum^n_{i=1} (x_i - \bar x)^2 }} {\sqrt {\sum^n_{i=1} (y_i - \bar y)^2 }} }
\end{equation}

```{r 225, message=FALSE, warning=FALSE}
c_m_corr_mat <- cor(c_m_RAW_AAC[, c(2, 4:23)], method = "p") # "p": Pearson test for continous variables

corrplot(abs(c_m_corr_mat),
         title = "Correlation Plot Of AAC, RAW Data",
         method = "square",
         type = "lower",
         tl.pos = "d",
         cl.lim = c(0, 1),
         addgrid.col = "lightgrey",
         cl.pos = "b",                   # Color legend position bottom.
         order = "FPC",                  # "FPC" = first principal component order.
         mar = c(1, 2, 1, 2),
         tl.col = "black")
```

NOTE: Amino acids shown in First Principal Component order, top to bottom.

1. Maximum value of Correlation between T & N.
```{r 226, echo=FALSE}
c_m_corr_mat["T", "N"]
```

2. According to Max Kuhn[^28], correlation coefficients need only be addressed if the |R| >= 0.75.
3. Therefore is **no reason to consider multicollinearity**.

[^28]:Max Kuhn and Kjell Johnson, Applied Predictive Modeling, Springer Publishing, 2018, P.47 (http://appliedpredictivemodeling.com/)

[^29]:Max Kuhn and Kjell Johnson, Applied Predictive Modeling, Springer Publishing, 2018, (http://appliedpredictivemodeling.com/)

### Boruta Random Forest Test, RAW data

>It finds relevant features by comparing original attributes' importance with importance achievable at random, estimated using their permuted copies (shadows).
>
>Miron Kursa [^211]

[^211]:https://notabug.org/mbq/Boruta/

```{r 227}
c_m_class_20 <- c_m_RAW_AAC[, -c(2, 3)] # Remove TotalAA & PID
Class <- as.factor(c_m_class_20$Class) # Convert 'Class' To Factor
```

NOTE: *mcAdj = TRUE*, If True, multiple comparisons will be adjusted using the Bonferroni method to calculate p-values. Therefore, $p_i \leq \large \frac {\alpha} {m}$ where $\alpha$ is the desired p-value and $m$ is the total number of null hypotheses. 
```{r 228, message=FALSE, warning=FALSE, cache=TRUE}
set.seed(1000)
#registerDoMC(cores = 3) # Start multi-processor mode
start_time <- Sys.time() # Start timer

boruta_output <- Boruta(Class ~ .,
                        data = c_m_class_20[, -1],
                        mcAdj = TRUE, # See Note above.
                        doTrace = 1) # doTrace = 1, represents non-verbose mode.

#registerDoSEQ() # Stop multi-processor mode
end_time <- Sys.time() # End timer
end_time - start_time # Display elapsed time
```

```{r 229, eval=FALSE}
names(boruta_output)
```

### Plot variable importance, RAW Data

```{r 230, echo=FALSE}
plot(boruta_output,
     cex.axis = 1,
     las = 2,
     ylim = c(-5, 50),
     main = "Variable Importance, RAW Data (Bigger=Better)")
```

### Variable importance scores, RAW Data

```{r 231, echo=FALSE, message=FALSE}
roughFixMod <- TentativeRoughFix(boruta_output)
imps <- attStats(roughFixMod)
imps2 <- imps[imps$decision != "Rejected", c("meanImp", "decision")]
meanImps <- imps2[order(-imps2$meanImp), ] # descending sort

kable(meanImps) %>% kable_styling(bootstrap_options = c("striped", "hover"))

# knitr::kable(meanImps,
# full_width = F,
# position = "left",
# caption = "Mean Importance Scores & Decision, RAW Data")
```

### Conclusion for Boruta random forest test, RAW Data

- All features are essential. None should be dropped.

### Conclusions For EDA, RAW data

Three amino acids (C, F, I) from the single amino acid percent composition were deemed problematic due to their skewness were greater than 2.0. This suggests that a transformation should be carried out to rectify this issue.

| Protein          | Skewness |
| :--------------- | :------: |
| C, Cysteine      | 2.538162 |
| F, Phenolalanine | 2.128118 |
| I, Isoleucine    | 2.192145 |

---

## Analysis of TRANSFORMED data

**This EDA section is a reevaluation square root transformed, `c_m_RAW_ACC.csv` data set, hence called `c_m_TRANSFORMED.csv.` **

The $\sqrt x_i$ *Transformed* data is derived from `c_m_RAW_ACC.csv` where the amino acids C, F, I were transformed using a square root function. This transformation was done to reduce the skewness of these samples and avoid modeling problems arising from high skewness, as seen below.

| Amino Acid       | Initial skewness | Skew After Square-Root Transformation |
| :--------------- | :--------------: | :-----------------------------------: |
| C, Cysteine      |     2.538162     |               0.3478132               |
| F, Phenolalanine |     2.128118     |               -0.102739               |
| I, Isoleucine    |     2.192145     |               0.2934749               |

```{r 232, message=FALSE, include=FALSE}
# Import libraries NO "doMC",
Libraries <- c("knitr", "readr", "RColorBrewer", "corrplot",  "Boruta")
for (i in Libraries) 
    {
    library(i, character.only = TRUE)
    }
opts_chunk$set(cache = TRUE, warning = FALSE, message = FALSE)
```

```{r 233, message=FALSE, warning=FALSE}
# Import Transformed data
c_m_TRANSFORMED <- read_csv("./00-data/02-aac_dpc_values/c_m_TRANSFORMED.csv")
Class <- as.factor(c_m_TRANSFORMED$Class)
```

### Check Transformed dataframe dimensions

```{r 234}
dim(c_m_TRANSFORMED)
```

### Check Transformed for missing values

```{r 235}
apply(is.na(c_m_TRANSFORMED), 2, which)
```

- No missing values found.

### Count Transformed data for the number of polypeptides per class

Number of polypeptides per Class:

- Class 0 = Control,
- Class 1 = Myoglobin
```{r 236, echo=FALSE}
class_table <- table(c_m_TRANSFORMED$Class)
class_table
```

### Visualization Descriptive Statistics, TRANSFORMED data

Formulas for mean:
\begin{equation} 
E[X] = \sum_{i=1}^n x_i p_i ~~; ~~~~~~ \bar x = \frac {1}{n} \sum_{i=1}^n x_i
\end{equation}

### Scatter plot of means of *Myoglobin-Control* amino acid composition $\sqrt x_i$, TRANSFORMED data

- This plot shows the means for each feature (column-means) in the dataset. The means represent the ungrouped or total of all proteins (where n=2340) versus AA type.

```{r 237, echo=FALSE}
AA_ave <- colMeans(c_m_TRANSFORMED[, 4:23])
plot(AA_ave,
     main = "Column-Means Vs Amino Acid, TRANSFORMED data",
     ylab = "% Composition",
     xlab = "Amino Acid",
     sub = "(Note: The red line at 0.1 is simply an arbitrary marker)",
     ylim = c(0, 0.3),
     type = "b",
     xaxt = "n")
axis(1, at = 1:20, labels = names(c_m_TRANSFORMED[, 4:23]))
```

```{r 238, echo=FALSE, cache=TRUE}
# * Pseudo-code for Data manipulation for Grouped Bar charts
# + A-1. Subset 7 protein groups, [Control:Ctrl, Myoglobin:Mgb] & Grand-Mean of both sets
# + A-2. Determine column means for each protein class
# + A-3. Calculate percentage values
# + A-4. Produce Grouped Bar Plot

# A-1
ctrl_set <- c_m_TRANSFORMED[which(c_m_TRANSFORMED$Class == "0"), ]
mgb_set <- c_m_TRANSFORMED[which(c_m_TRANSFORMED$Class == "1"), ]

# A-2
ctrl_means <- apply(ctrl_set[, 4:23], 2, mean)
mgb_means  <- apply(mgb_set[, 4:23], 2, mean)
grand_mean <- apply(c_m_TRANSFORMED[, 4:23], 2, mean)

# A-3
data <- data.frame(ctrl_means, mgb_means, grand_mean)
percent_aa <- as.matrix(t(100 * data))
```

```{r 239, eval=FALSE}
# A-4
## Grouped barchart of $\sqrt x_i$ Transformed amino acid vs.
## protein category data
barplot(percent_aa,
        main = "Mean % A.A.Composition, TRANSFORMED data",
        ylab = "% AA Composition",
        ylim = c(0, 30),
        col = colorRampPalette(brewer.pal(4, "Blues"))(3),
        legend = T,
        beside = T)
```

### Grouped bar chart of means for percent amino acid composition of Transformed Data; control & myoglobin categories

```{r 240, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
data2 <- data.frame(ctrl_means, mgb_means)
percent_aa2 <- as.matrix(t(100 * data2))

barplot(percent_aa2,
        ylim = c(0, 30),
        main = "Means of % A.A.Composition, TRANSFORMED data",
        ylab = "% AA Composition",
        col = colorRampPalette(brewer.pal(4, "Blues"))(2),
        legend = T,
        beside = T)
```

### Boxplots of grand-means of the overall amino acid composition of square-root transformed data

```{r 241, echo=FALSE, message=FALSE, warning=FALSE}
boxplot(c_m_TRANSFORMED[, 4:23],
        main = "% AAC Vs Amino Acid, TRANSFORMED data",
        ylab = "% AAC",
        xlab = "Amino Acid",
        las = 1)
```

### Boxplots of amino acid compositions for control (only) of square-root transformed data

```{r 242, echo=FALSE}
boxplot(ctrl_set[, 4:23],
        main = "Control, % AAC Vs Amino Acid, TRANSFORMED data",
        ylab = "% AAC",
        xlab = "Amino Acid",
        las = 1)
```

### Boxplots of amino acid compositions for myoglobin of square-root transformed Data(only), TRANSFORMED data

```{r 243, echo=FALSE}
boxplot(mgb_set[, 4:23],
        main = "Myoglobin, % AAC Vs Amino Acid, TRANSFORMED data",
        ylab = "% AAC",
        xlab = "Amino Acid",
        las = 1)
```

### Boxplots Of Length Of Polypeptides Of Transformed Data; Myoglobin, Control & Combined

```{r 244, echo=FALSE}
ctrl_totalaa  <- ctrl_set[, 2]
mgb_totalaa   <- mgb_set[, 2]
grand_totalaa <- c_m_TRANSFORMED[, 2]

data <- c(ctrl_totalaa, mgb_totalaa, grand_totalaa)

boxplot(data,
        ylim = c(0, 5000),
        main = "Length of Polypeptides, TRANSFORMED data",
        ylab = "Length of Polypeptides",
        xaxt = "n",
        las = 2)
axis(1, at = 1:3, labels = c("Control", "Myoglobin", "Combined"))
```



### Coefficient of Variance (CV), TRANSFORMED data

Standard deviations are sensitive to scale. Therefore I compare the normalized standard deviations. This normalized standard deviation is more commonly called the coefficient of variation (CV).

\begin{equation} 
CV = \frac {\sigma (x)} {E [|x|]} ~~~ where ~~~ \sigma(x) \equiv \sqrt{ E[x - \mu]^2 }
\end{equation}

\begin{equation} 
CV ~~=~~ \frac{1}{\bar x} \cdot \sqrt{ \frac{1}{n-1} \sum^n_{i=1} (x_i - \bar x)^2}
\end{equation}

### Plot of Coefficient Of Variance (CV)

```{r 245, echo=FALSE}
AA_var_norm <- (apply(c_m_TRANSFORMED[, 4:23], 2, sd)) / AA_ave
plot(AA_var_norm,
     main = "Coefficient Of Variance, TRANSFORMED data",
     ylab = "Coefficient Of Variance (CV)",
     xlab = "Amino Acid",
     ylim = c(0, 1.5),
     type = "b",
     xaxt = "n")
axis(1, at = 1:20, labels = names(c_m_TRANSFORMED[, 4:23]))
```
```{r 246}
AA_var_norm
```



### Skewness of distributions, TRANSFORMED data

\begin{equation} 
Skewness ~= E\left[ \left( \frac{X - \mu}{\sigma(x)} \right)^3 \right] ~~~~ where ~~~~ \sigma(x) \equiv \sqrt{ E[x - \mu]^2 }
\end{equation}

\begin{equation} 
Skewness ~= \frac { \frac{1}{n} \sum^n_{i=1} (x_i - \bar x)^3 } { \left( \sqrt{ \frac{1}{n-1} \sum^n_{i=1} (x_i - \bar x)^2 } \right) ^ {3}}
\end{equation}

- Skewness values for each A.A. by Class of square-root transformed data

```{r 247, echo=FALSE}
AA_skewness <- (apply(c_m_TRANSFORMED[, 4:23], 2, e1071::skewness))
plot(AA_skewness,
     main = "Skewness of Amino Acids, TRANSFORMED data",
     sub = "(Red bar at Skew=2.0 indicates all values good.)",
     ylab = "Skewness",
     xlab = "Amino Acid",
     type = "b",
     ylim = c(-0.5, 2.3),
     xaxt = "n")
axis(1, at = 1:20, labels = names(c_m_TRANSFORMED[, 4:23]))
abline(h = 2.0, col = "red")
```

```{r 248}
AA_skewness
```

### Determine coefficients of correlation, TRANSFORMED data

An easily interpretable test is a correlation 2D-plot for investigating multicollinearity or feature reduction. Fewer attributes "means decreased computational time and complexity. Secondly, if two predictors are highly correlated, this implies that they measure the same underlying information. Removing one should not compromise the performance of the model and might lead to a more parsimonious and interpretable model. Third, some models can be crippled by predictors with degenerate distributions." [^210]

[^210]:Max Kuhn and Kjell Johnson, Applied Predictive Modeling, Springer Publishing, 2018

Pearson's correlation coefficient: 
\begin{equation} 
\rho_{x,y} = \frac {E \left[(X - \mu_x)(X - \mu_y) \right]} {\sigma_x \sigma_y}
\end{equation}

\begin{equation} 
r_{xy} = \frac {\sum^n_{i=1} (x_i - \bar x)(y_1 - \bar y)} { {\sqrt {\sum^n_{i=1} (x_i - \bar x)^2 }} {\sqrt {\sum^n_{i=1} (y_i - \bar y)^2 }} }
\end{equation}

```{r 250, echo=FALSE}
c_m_corr_mat <- cor(c_m_TRANSFORMED[, c(2, 4:23)],
                    method = "p") # "p": Pearson test for continous variables

corrplot(abs(c_m_corr_mat),
         title = "Correlation Plot, TRANSFORMED data",
         method = "square",
         type = "lower",
         tl.pos = "d",
         cl.lim = c(0, 1),
         addgrid.col = "lightgrey",
         cl.pos = "b", # Color legend position bottom.
         order = "FPC", # "FPC" = first principal component order.
         mar = c(1, 2, 1, 2),
         tl.col = "black")
```

```{r 251}
c_m_corr_mat["T", "N"]
```

**No values in the correlation matrix meet the 0.75 cut off criteria for problems.**

### Boruta - Dimensionality Reduction, TRANSFORMED data

```{r 252, echo=FALSE}
c_m_class_20 <- c_m_TRANSFORMED[, -c(2, 3)] # Remove TotalAA & PID
Class <- as.factor(c_m_class_20$Class) # Convert 'Class' To Factor
```

**Perform Boruta search**

NOTE: *mcAdj = TRUE*: If True, multiple comparisons will be adjusted using the Bonferroni method to calculate p-values. Therefore, $p_i \leq \frac {\alpha} {m}$ where $\alpha$ is the desired p-value and $m$ is the total number of null hypotheses. 
```{r 253, cache=TRUE}
set.seed(1000)
#registerDoMC(cores = 3) # Start multi-processor mode
start_time <- Sys.time() # Start timer

boruta_output <- Boruta(Class ~ .,
                        data = c_m_class_20[, -1],
                        mcAdj = TRUE, # See Note above.
                        doTrace = 1) # doTrace = 1, represents non-verbose mode.

#registerDoSEQ() # Stop multi-processor mode
end_time <- Sys.time() # End timer
end_time - start_time # Display elapsed time
```

### Plot Variable Importance, TRANSFORMED data

```{r 254}
plot(boruta_output,
     cex.axis = 1,
     las = 2,
     ylim = c(-5, 50),
     main = "Variable Importance, TRANSFORMED data(Bigger=Better)")
```

### Variable Importance Scores, TRANSFORMED data

```{r 255, message=FALSE}
roughFixMod <- TentativeRoughFix(boruta_output)
imps <- attStats(roughFixMod)
imps2 <- imps[imps$decision != "Rejected", c("meanImp", "decision")]
meanImps <- imps2[order(-imps2$meanImp), ] # descending sort

kable(meanImps) %>% kable_styling(bootstrap_options = c("striped", "hover"))

# knitr::kable(meanImps,
# full_width = F,
# position = "left",
# caption = "Mean Importance Scores & Decision, TRANSFORMED data")
```

```{r 256, eval=FALSE, include=FALSE}
## Plot importance history
plotImpHistory(boruta_output)
```

The *Boruta Random Rorest test* shows that all features are essential therefore none should be dropped from TRANSFORMED data.

