---
title: "11-pca-appendix"
author: "mcc"
date: "4/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r 33, message=FALSE, warning=FALSE, include=FALSE}
# Import libraries
library(easypackages)
libraries <- c("knitr", "ggfortify")
```

## Principle component analysis using `norm_c_m_20aa`

```{r 35, cache=TRUE}
start_time <- Sys.time() # Start timer

c_m_20_PCA <- prcomp(norm_c_m_20aa)

Sys.time() - start_time # End timer & display time difference
```

#### Biplot 1: PC1 Vs. PC2 with 'Class' by color labels

- Black indicates control protein set, Class = 0

- Blue indicates myoglobin protein set, Class = 1
```{r 38,eval=FALSE, fig.align="center", fig.height=6, fig.width=6, include=FALSE, cache=TRUE}
autoplot(c_m_20_PCA,
         data = c_m_transformed,
         label.show.legend = F,
         colour = "Class")
         # shape = FALSE,
         # label = TRUE,
         # label.size = 4,
         # loadings = TRUE, 
         # loadings.label = TRUE, 
         # loadings.label.size = 7,
         # loadings.label.hjust = 2)
```

The first two principal components describe 46.95% of the variance.

```{r 39,eval=FALSE, include=FALSE, cache=TRUE}
# Print Biplot1.annotated.png
# Biplot 1 for Conclusion discussion
png(filename = "Biplot1.annotated.png",
    width = 6,
    height = 6,
    units = "in",
    res = 300)
    autoplot(c_m_20_PCA,
    data = c_m_transformed,
    label.show.legend = F,
    colour = "Class",
    shape = FALSE,
    label = TRUE,
    label.size = 4,
    loadings = TRUE,
    loadings.label = TRUE,
    loadings.label.size = 7,
    loadings.label.hjust = 2)
dev.off()
```

#### Biplot 2: Determination Of 4 Rule Set For Outliers

```{r 310,echo=FALSE, fig.height=6, fig.width=6, fig.align="center", cache=TRUE}
plot(x = c_m_20_PCA$x[, 1],
     y = c_m_20_PCA$x[, 2],
     main = "Boundary (Outlier) Determination of PC1 Vs PC2",
     xlab = "PC2",
     ylab = "PC1",
     cex = 0.5)
# col = c("#bbbbbb", "#ff0000"),
# cols <- c(1,2))
abline(v = 3, col = "blue", lw = 1)
abline(v = -3, col = "blue", lw = 1)
abline(h = -3, col = "blue", lw = 1)
abline(h = 3, col = "blue", lw = 1)
```

```{r 311,eval=FALSE, include=FALSE, cache=TRUE}
# Print 'Biplot 2'. SAVE FOR LATER USE.
# Biplot 2 for Conclusion discussion
png(filename = "./00-data/10-images/Biplot.2.png",
    width = 6,
    height = 6,
    units = "in",
    res = 300)
plot(x = c_m_20_PCA$x[, 1],
     y = c_m_20_PCA$x[, 2],
     main = "Boundary (Outlier) Determination of PC1 Vs PC2",
     xlab = "PC2",
     ylab = "PC1",
     cex = 0.5)
# col = c("#bbbbbb", "#ff0000"),
# cols <- c(1,2))
abline(v = 3, col = "blue", lw = 1)
abline(v = -3, col = "blue", lw = 1)
abline(h = -3, col = "blue", lw = 1)
abline(h = 3, col = "blue", lw = 1)
dev.off()
```


### List of all outliers (union and sorted) found using the ruleset 1 through 4

- The list of total outliers is derived by taking the `union` of `outliers_PC1` and `outliers_PC2` and then using `sort.`

```{r 312,cache=TRUE}
total_pca_1_2_outliers <- union(outliers_PC1, outliers_PC2)
total_pca_1_2_outliers <- sort(total_pca_1_2_outliers)

length(total_pca_1_2_outliers)

# Write out to Outliers folder
write.table(total_pca_1_2_outliers,
file = "./00-data/03-ml_results/pca_outliers.csv",
row.names = FALSE,
na = "",
col.names = "rowNum",
sep = ",")
```

