---
title: "14-svm-appendix"
author: "mcc"
date: "4/17/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## SVM-Linear Model

```{r 610, include=FALSE}
opts_chunk$set(cache = TRUE, fig.align = "center")
```

```{r 611}
# Import data & data handling
c_m_TRANSFORMED <- read_csv("./00-data/02-aac_dpc_values/c_m_TRANSFORMED.csv",
                            col_types = cols(Class = col_factor(levels = c("0", "1")),
                                             PID = col_skip(),
                                             TotalAA = col_skip()))
```

Partition data into training and testing sets
```{r 612}
set.seed(1000)
index <- createDataPartition(c_m_TRANSFORMED$Class, p = 0.8, list = FALSE)

training_set <- c_m_TRANSFORMED[ index,]
test_set     <- c_m_TRANSFORMED[-index,]

Class_test <- as.factor(test_set$Class)
```

### SVM-Linear Training

```{r 613, cache=TRUE}
set.seed(1000)
start_time <- Sys.time() # Start timer

# tuneGrid = svmLinearGrid
svmLinearGrid <- expand.grid(C = c(2^(4.5), 2^5, 2^(5.5)))

# Create model, 10X fold CV repeated 5X
tcontrol <- trainControl(method = "repeatedcv",
                         number = 10,
                         repeats = 5,
                         savePredictions = "final") # Saves predictions

lin_model_obj <- train(Class ~ .,
                       data = training_set,
                       method = "svmLinear",
                       trControl = tcontrol,
                       tuneGrid = svmLinearGrid)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```

### SVM-Linear Model Summary

```{r 614}
lin_model_obj
```

### SVM-Linear Predict test_set

```{r 615, cache=TRUE}
Predicted_test_vals <- predict(lin_model_obj, test_set[, -1])

summary(Predicted_test_vals)
```

### SVM-Linear Confusion Matrix

```{r 616}
confusionMatrix(Predicted_test_vals, Class_test, positive = "1")
```

### SVM-Linear Obtain False Positives & False Negatives

```{r 617}
fp_fn_svm_linear <- lin_model_obj %>% pluck("pred") %>% dplyr::filter(obs != pred)

# Write out to Outliers folder
write.table(fp_fn_svm_linear,
            file = "./00-data/03-ml_results/fp_fn_svm_linear.csv",
            row.names = FALSE,
            na = "",
            col.names = TRUE,
            sep = ",")

nrow(fp_fn_svm_linear)
head(fp_fn_svm_linear)
```

---

## SVM-Polynomial Model

Partition data into training and testing sets
```{r 618}
set.seed(1000)
index <- createDataPartition(c_m_TRANSFORMED$Class, p = 0.8, list = FALSE)

training_set <- c_m_TRANSFORMED[ index,]
test_set     <- c_m_TRANSFORMED[-index,]

Class_test <- as.factor(test_set$Class)
```

### SVM-Poly Training

```{r 619, cache=TRUE}
set.seed(1000)
start_time <- Sys.time() # Start timer

# Create model, 10X fold CV repeated 5X
tcontrol <- trainControl(method = "repeatedcv",
                         number = 10,
                         repeats = 5,
                         savePredictions = "final") # Saves predictions

poly_model_obj <- train(Class ~ .,
                        data = training_set,
                        method = "svmPoly",
                        trControl= tcontrol)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```

### SVM-Poly Model Summary

```{r 620}
poly_model_obj
```

### SVM-Poly Predict test_set
```{r 621, cache=TRUE}
Predicted_test_vals <- predict(poly_model_obj, test_set[, -1])

summary(Predicted_test_vals)
```

### SVM-Poly Confusion Matrix
```{r 622}
confusionMatrix(Predicted_test_vals, Class_test, positive = "1")
```

### SVM-Poly Obtain False Positives & False Negatives

```{r 623}
fp_fn_svm_poly <- poly_model_obj %>% pluck("pred") %>% dplyr::filter(obs != pred)

# Write CSV in R
write.table(fp_fn_svm_poly, 
            file = "./00-data/03-ml_results/fp_fn_svm_poly.csv", 
            row.names = FALSE, 
            na = "", 
            col.names = TRUE, 
            sep=",")

nrow(fp_fn_svm_poly)
head(fp_fn_svm_poly)
```

## SVM-RBF Model

Import data & data handling
```{r 625}
c_m_TRANSFORMED <- read_csv("./00-data/02-aac_dpc_values/c_m_TRANSFORMED.csv",
                            col_types = cols(Class = col_factor(levels = c("0","1")), 
                                             PID = col_skip(), TotalAA = col_skip()))
#View(c_m_TRANSFORMED)
```

Partition data into training and testing sets
```{r 626}
set.seed(1000)
index <- createDataPartition(c_m_TRANSFORMED$Class, p = 0.8, list = FALSE)

training_set <- c_m_TRANSFORMED[ index,]
test_set     <- c_m_TRANSFORMED[-index,]

Class_test <- as.factor(test_set$Class)
```

### SVM-RBF Training
```{r 627, cache=TRUE}
set.seed(1000)
start_time <- Sys.time() # Start timer

# Create tuneGrid: Cost
tune.Grid = data.frame(expand.grid(C = 2^(seq(-5, 15, 2))))
                                          
# Create: 10X fold CV repeated 5X
tcontrol <- trainControl(method = "repeatedcv",
                         number = 10,
                         repeats = 5,
                         savePredictions = "final") # Save predictions

rbf_model_obj <- train(Class ~ .,
                   data = training_set,
                   method = "svmRadialCost",
                   tuneGrid = tune.Grid,
                   trControl= tcontrol)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
```

### SVM-RBF Model Summary
```{r 628}
rbf_model_obj
```

### SVM-RBF Predict test_set
```{r 629, cache=TRUE}
Predicted_test_vals <- predict(rbf_model_obj, test_set[, -1])

summary(Predicted_test_vals)
```

### SVM-RBF Confusion Matrix
```{r 630}
confusionMatrix(Predicted_test_vals, Class_test, positive = "1")
```

### SVM-RBF Obtain False Positives & False Negatives
```{r 631}
fp_fn_svmRadialCost <- rbf_model_obj %>% pluck("pred") %>% dplyr::filter(obs != pred)

# Write CSV in R
write.table(fp_fn_svmRadialCost, 
            file = "./00-data/03-ml_results/fp_fn_svmRbf.csv", 
            row.names = FALSE, 
            na = "", 
            col.names = TRUE, 
            sep=",")

nrow(fp_fn_svmRadialCost)
head(fp_fn_svmRadialCost)
```
