---
title: "R Notebook Title"
output: html_notebook
---
Matthew Curcio
Worcester Polytechnic Inst.
Class: Machine Learning wrt Bioinformatics

Outline:
1. Intro.
  1.1 mention paper on SVM and O2 binding proteins, thesis and ideas...
  1.2 Using R package: Caret
2. Data
  2.1 How many proteins? What type of proteins? 
      How to choose random proteins from PDB(?)
  2.2 Transform data into numerical format, monopeptide, dipeptides
  2.3 Clean data
  2.4 Investigate date using
    2.4.1 Trellis
    2.4.2 Histograms
    2.4.3 boxplots
    2.4.4 Summary of data min, max, mean, ??
3. Data Analysis
  3.1 Cros-validation for model performance estimation
  3.2 See Example: https://rpubs.com/lilsummer1989/185374
  3.3 
    
  
  
  
References:
1. https://rpubs.com/rpadebet/269829

2. http://rischanlab.github.io/RandomForest.html

3. https://www.youtube.com/watch?v=TfVuakuPcpE&t=1s&index=42&list=WL

Test of 'rpart' using the Iris data set from R.Fisher.

```{r}
library(rpart)
library(rpart.plot)
```

```{r}
library(datasets)
data(iris)
dim(iris)

plot(iris)

summary(iris)
```
```{r}
# use ?plot to read more about other arguments
?plot()
```


```{r}
# Load the Caret package which allows us to partition the data
library(caret)
# We use the dataset to create a partition (80% training 20% testing)
index <- createDataPartition(iris$Species, p=0.80, list=FALSE)
# select 20% of the data for testing
testset <- iris[-index,]
# select 80% of data to train the models
trainset <- iris[index,]

dtm <- rpart(Species~., data = trainset, method = "class")

```

```{r}
dtm
```

```{r}
rpart.plot(dtm, type = 0, extra = 102)

```

```{r}
rpart.plot(dtm, type = 1, extra = 102)
```

```{r}
rpart.plot(dtm, type = 2, extra = 102)
```

```{r}
rpart.plot(dtm, type = 3, extra = 102)
```

```{r}
rpart.plot(dtm, type = 4, extra = 102)
```

```{r}
## Box plot to understand how the distribution varies by class of flower
par(mfrow=c(1,4))
for(i in 1:4) {
    boxplot(trainset[,i], main = names(trainset)[i])
}
```

```{r}
# Begin using GGplot2
library(ggplot2)

# Scatter plot
# g <- ggplot(data=trainset, aes(x = Petal.Length, y = Petal.Width))
# print(g)

g <- g + 
    geom_point(aes(color=Species, shape=Species)) +
    xlab("Petal Length") +
    ylab("Petal Width") +
    ggtitle("Petal Length-Width")+
    geom_smooth(method="lm")
print(g)
```

```{r}
## Box Plot
box <- ggplot(data=trainset, aes(x=Species, y=Sepal.Length)) +
    geom_boxplot(aes(fill=Species)) + 
    ylab("Sepal Length") +
    ggtitle("Iris Boxplot") +
    stat_summary(fun.y=mean, geom="point", shape=5, size=4) 
print(box)

```

```{r}
library(ggthemes)
## Histogram
histogram <- ggplot(data=iris, aes(x=Sepal.Width)) +
    geom_histogram(binwidth=0.2, color="black", aes(fill=Species)) + 
    xlab("Sepal Width") +  
    ylab("Frequency") + 
    ggtitle("Histogram of Sepal Width")+
    theme_economist()
print(histogram)
```

```{r}
## Faceting: Producing multiple charts in one plot
library(ggthemes)
facet <- ggplot(data=trainset, aes(Sepal.Length, y=Sepal.Width, color=Species))+
    geom_point(aes(shape=Species), size=1.5) + 
    geom_smooth(method="lm") +
    xlab("Sepal Length") +
    ylab("Sepal Width") +
    ggtitle("Faceting") +
    theme_fivethirtyeight() +
    facet_grid(. ~ Species) # Along rows
print(facet)
```

```{r}
print(dtm)
```


```{r}
## Predictions on train dataset
pred <- table(predict(object = dtm, 
                      newdata = trainset[,1:4],
                      type="class"))
pred
```

```{r}
confusionMatrix(
```






