---
title: "pca_text_label_plot"
author: "MCC"
date: "12/23/2019"
output: html_document
---

First, let’s use the plot() function to plot the points.

##### Labelling Points in a Scatter Plot

plot(sr~dpi, xlim = c(0, 3500), xlab = 'Real Per-Capita Disposable Income', ylab = 'Aggregate Personal Savings', main = 'Intercountry Life-Cycle Savings Data', data = LifeCycleSavings[1:9,])

Then, let’s use the text() function to add the text labels to the data.  It has to be nested within the with() function, because, unlike plot(), “data” is not a valid option for text().

with(LifeCycleSavings[1:9,], text(sr~dpi, labels = row.names(LifeCycleSavings[1:9,]), pos = 4))

```{r}
c_m_TRANSFORMED <- read.csv("~/Dropbox/a1_mcc_project/00-data/02-aac_dpc_values/c_m_TRANSFORMED.csv",
                            stringsAsFactors=FALSE)
#View(c_m_TRANSFORMED)

c_m_transformed <- c_m_TRANSFORMED[,-c(1:3)]

pca_vals <- prcomp(c_m_transformed, scale = TRUE)

PC1_2_values <- pca_vals$rotation[,1:2]
#PC1_2_values

```

First, let’s use the plot() function to plot the points.

##### Labelling Points in a Scatter Plot

plot(sr~dpi, xlim = c(0, 3500), xlab = 'Real Per-Capita Disposable Income', ylab = 'Aggregate Personal Savings', main = 'Intercountry Life-Cycle Savings Data', data = LifeCycleSavings[1:9,])

Then, let’s use the text() function to add the text labels to the data.  It has to be nested within the with() function, because, unlike plot(), “data” is not a valid option for text().

with(LifeCycleSavings[1:9,], text(sr~dpi, labels = row.names(LifeCycleSavings[1:9,]), pos = 4))

============================
plot(0,0) ## start a plot
opar <- par(new = TRUE, ## add a new layer
            mar = c(0,0,0,0)) ## with no margins margins
## set up the plotting area for this layer
plot(1,1,xlim=c(0,1),ylim=c(0,1),type='n',xlab='',ylab='') 
# arrows(0.6,0.6,0.8,1) ## add arrow
par(opar) ## return the plot parameters to their prior values
=======================================================


```{r fig.height=5, fig.width=5}
plot(PC2~PC1, 
     xlim = c(-0.5, 0.5), 
     ylim = c(-0.5, 0.5),
     xlab = 'PC1', 
     ylab = 'PC2', 
     main = 'PCA Loadings Vs Outliers', 
     data = pca_vals$rotation[,1:2])
plot(0,0,
     xlim = c(-0.5, 0.5), 
     ylim = c(-0.5, 0.5))
with(pca_vals, text(PC2 ~ PC1,
                    labels = row.names(pca_vals$rotation),
                    pos = 3))

```

```{r}
pca_values <- prcomp(c_m_TRANSFORMED, scale = TRUE)
biplot(pca_values)
#summary(pca_values)
```