# Support Vector Machine Radial Basis Function, Using `kernlab`

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      cache.lazy = FALSE)
```

## Summary

NOTE: SVM {caret} / https://rpubs.com/PranovMishra/476455

## Pseudocode



## Hyper-parameters

## Big O

## Model Training & Tuning

Libraries
```{r, message=FALSE, warning=FALSE}
Libraries = c("doMC", "caret", "kernlab")

for(p in Libraries){  # Install if not present
    if(!require(p, character.only = TRUE)) { install.packages(p, dependencies = TRUE) }
    library(p, character.only = TRUE)
}
```

Import data & data handling

```{r}
test_harness_paa <- read.csv("data/test_harness_paa.csv")
test_harness_paa <- test_harness_paa[, -c(2,3)]
Class <- as.factor(test_harness_paa$Class) # Convert ‘Class’ To Factor
```

Partition data - training / testing sets

```{r}
set.seed(1000)
index <- createDataPartition(test_harness_paa$Class, p = 0.8, list = FALSE)

training_set <- test_harness_paa[ index,]
test_set     <- test_harness_paa[-index,]

Class_test <- as.factor(test_set$Class)
```

### Run Training Set 
 
```{r cache=TRUE}
set.seed(1000)
registerDoMC(cores = 3) # Start multi-processor mode
start_time <- Sys.time() # Start timer

# Create grid control: sigma, C
# tune.Grid = data.frame(expand.grid(sigma = seq(1, 10, 1)),
#                                    C = seq(1, 10, 1))
                                          
# Create model, 10X fold CV repeated 5X
tcontrol <- trainControl(method = "repeatedcv",
                         number = 5,
                         repeats = 3)

model_list <- train(Class ~ .,
                    data = training_set,
                    methodList = "svmRadial",
                    #tuneGrid = tune.Grid,
                    trControl = tcontrol)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
registerDoSEQ() # Stop multi-processor mode
```

### Display Model Results

Model summary

```{r}
model_list
```

## Predictions - Run Test Set

Predict test set
 
```{r}
Predicted_test_vals <- predict(model_list, test_set[,-1])

length(Predicted_test_vals)
```

Quick summary

```{r}
summary(Predicted_test_vals)
```

Inspect Model Object

```{r, cache=TRUE}
names(model_list)
```

Confusion matrix

```{r}
confusionMatrix(Predicted_test_vals, Class_test)
```



