# Support Vector Machine Linear Kernel, Using `kernlab`

```{r, include=FALSE}
library(knitr)
knitr::opts_chunk$set(cache = TRUE, 
                      warning = FALSE, 
                      message = FALSE, 
                      cache.lazy = FALSE)
```

## Summary

NOTE: SVM {caret} / https://rpubs.com/PranovMishra/476455

## Pseudocode



## Hyper-parameters

## Big O

## Model Training & Tuning

Libraries
```{r, message=FALSE, warning=FALSE}
Libraries = c("doMC", "caret", "kernlab")

for(p in Libraries){  # Install if not present
    if(!require(p, character.only = TRUE)) { install.packages(p, dependencies = TRUE) }
    library(p, character.only = TRUE)
}
```

Import data & data handling

```{r}
test_harness_paa <- read.csv("data/test_harness_paa.csv")
test_harness_paa <- test_harness_paa[, -c(2,3)]
Class <- as.factor(test_harness_paa$Class) # Convert ‘Class’ To Factor
```

Partition data - training / testing sets
```{r}
set.seed(1000)
index <- createDataPartition(test_harness_paa$Class, p = 0.8, list = FALSE)

training_set <- test_harness_paa[ index,]
test_set     <- test_harness_paa[-index,]

Class_test <- as.factor(test_set$Class)
```

### Run Training Set 
 
```{r cache=TRUE}
set.seed(1000)
registerDoMC(cores = 3) # Start multi-processor mode
start_time <- Sys.time() # Start timer

# Tuning parameter C = cost for optimized model
# grid = expand.grid(C = seq(0.5, 10, 0.5)) USE 2^-10 to 2^15 ????? SEE PAPER

# Create model, 10X fold CV repeated 5X
train_control <- trainControl(method = "repeatedcv",
                              number = 5,
                              repeats = 5)
model_list = train(Class ~ ., 
                   data = training_set,
                   method = 'svmLinear', # Build SVM-Linear
                   trControl = train_control)
                   #preProcess = c("pca","scale","center"),
                   #tuneGrid = grid)

end_time <- Sys.time()   # End timer
end_time - start_time    # Display time
registerDoSEQ() # Stop multi-processor mode
```

### Display Model Results

Model summary

```{r}
model_list
```

## Predictions - Run Test Set

Predict test set
 
```{r}
Predicted_test_vals <- predict(model_list, test_set[,-1])

length(Predicted_test_vals)
```

Quick summary

```{r}
summary(Predicted_test_vals)
```

Plot Model Vs Cost

```{r, cache=TRUE}
# plot(model_list)
```

Inspect Model Object

```{r, cache=TRUE}
names(model_list)
```

Confusion matrix

```{r}
confusionMatrix(Predicted_test_vals, Class_test)
```



